---
title: "Installation and Motivating Example"
author: "borangao"
date: "2022-11-09"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Installation of MESuSiE

```{r,message=FALSE,warning=FALSE}
library(devtools)
# Install MESuSiE
install_github("borangao/MESuSiE",dependencies = FALSE)
# Load MESuSiE
library(MESuSiE)
```

# Motivating Example

The motivating example is based on a toy dataset outlined in the manuscript.

## 1. Data Characteristics

The dataset contains:

- Six SNPs categorized into three distinct LD clusters.
- Within each cluster, the SNPs show strong correlations.
- Between clusters, SNP correlations are weak.

### SNP Grouping

1. **Cluster 1**: Contains SNP 1 and 2.
2. **Cluster 2**: Contains SNP 3 and 4.
3. **Cluster 3**: Contains SNP 5 and 6.

### Causal SNPs: 

Based on our simulations, the SNPs are postulated to have the following causal relationships:

- **SNP 4**: Simulated as a shared causal SNP across ancestries.
- **SNP 2**: Simulated as the causal SNP specific to the BB ancestry.
- **SNP 6**: Simulated as the causal SNP specific to the WB ancestry.


## 2. Input of MESuSiE
The data is provided with the package. For the input of the MESuSiE, we require a list of summary statistics and a list of LD matrices from multiple ancestries.
```{r,message=FALSE,warning=FALSE}
data(summ_stat_list)
data(LD_list)
summ_stat_list
```
Each element of the summary statistics list is a data frame of summary statistics from each ancestry with column name SNP, Beta, Se, Z, and N representing SNP information, marginal effect size, standard error, Z-scores and number of sample.
```{r,message=FALSE,warning=FALSE}
LD_list
```
Each element of the LD matrices list is a matrix of LD from each ancestry, with column name being SNP name matched up with the summary statistics. 

## 3. Analysis 

### MESuSiE
```{r,message=FALSE,warning=FALSE}
MESuSiE_res<-meSuSie_core(LD_list,summ_stat_list,L=10)
MESuSiE_res$pip_config
```
We observed that three SNPs have a posterior inclusion probability (PIP) exceeding the 0.5 threshold. Upon further examination of the PIP for ancestry-specificity and shared traits, we identified that SNP4 is shared, while SNP2 and SNP6 are ancestry-specific. The categories within the credible set indicate the ancestries affected by the SNPs present in the set.

### SuSiE
```{r,message=FALSE,warning=FALSE}
library(susieR)
susie_WB<-susie_rss(summ_stat_list$WB$Z,LD_list$WB)
susie_BB<-susie_rss(summ_stat_list$BB$Z,LD_list$BB)
susie_WB$pip
susie_BB$pip
```
After executing the univariate SuSiE analysis, we determined that SNP 3 and 6 are signals specific to Europeans, while SNP 2 and 4 emerged as signals specific to Africans, using a PIP threshold of 0.5.

### Paintor
```{r,message=FALSE,warning=FALSE}
##We load the Paintor result directly
paintor_res<-read.table("/net/fantasia/home/borang/Susie_Mult/Revision_Round_1/Simulation/091223/data/toy_example/result/fig1.mcmc.paintor",header=T)
paintor_res$Posterior_Prob
```

Paintor identifies SNP2-5 as signals without distinguishing ancestry-specific or shared causal variant.


## 4. Visualization
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.width=8}
library(ggplot2)
library(dplyr)
###########################################################################
#
#
#             Result Plot Part
#
#
##########################################################################

custom_theme <- function() {
  theme(
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),  
    axis.title.x = element_text(size = 7, face="bold"),
    axis.title.y = element_text(size = 7, face="bold"),
    strip.text.x = element_text(size = 5),
    strip.text.y = element_text(size = 5),
    strip.background = element_blank(),
    legend.text = element_text(size=7),
    legend.title = element_text(size=7, face="bold"),
    plot.title = element_text(size=7, hjust = 0.5),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(), 
    axis.line = element_line(color = "black")
  )
}


#############################################################
#
#
#
#				Beta Plot
#
#
#
###############################################################
beta_WB_all<-c(0,0,0,0.02,0,0.02)
beta_BB_all<-c(0,0.02,0,0.02,0,0)
WB_bar_data <- data.frame(SNP=paste0("SNP",seq(1,6,1)),Effect = beta_WB_all,Signal =c("None","BB","None","Shared","None","WB"))
BB_bar_data <- data.frame(SNP=paste0("SNP",seq(1,6,1)),Effect = beta_BB_all,Signal =c("None","BB","None","Shared","None","WB"))
# Grouped
bar_WB<-ggplot(WB_bar_data, aes( y=Effect, x=SNP,fill = Signal)) + xlab("")+ylab("Effect")+
  geom_bar(position="dodge", stat="identity")+theme_bw()+scale_fill_manual("Signal",values=c("WB"="#F2C1B6","BB"="#ABDB9F","Shared" = "#6162B0","None"="gray"))+theme(legend.position="none")
bar_WB = bar_WB + custom_theme()
bar_BB<-ggplot(BB_bar_data, aes( y=Effect, x=SNP,fill = Signal)) + xlab("")+ylab("Effect")+
  geom_bar(position="dodge", stat="identity")+theme_bw()+scale_fill_manual("Signal",values=c("WB"="#F2C1B6","BB"="#ABDB9F","Shared" = "#6162B0","None"="gray"))+theme(legend.position="none")
bar_BB = bar_BB + custom_theme()
#############################################################
#
#
#
#				GWAS Plot
#
#
#
###############################################################
z_WB<-summ_stat_list$WB$Z
z_BB<-summ_stat_list$BB$Z
gwas_data<-data.frame("SNP"=paste0("SNP",seq(1,6,1)),"WB_Pvalue"= -log10(2*pnorm(-abs(z_WB))),"BB_Pvalue"=-log10(2*pnorm(-abs(z_BB))))
locuszoom_WB<-ggplot(gwas_data,aes(x=SNP,y=WB_Pvalue))+theme_bw()+geom_point(size =2)+
                       geom_hline(yintercept=8, linetype="dashed", color = "red")+
                       labs(x="",y="-log10 P-value") +custom_theme()
locuszoom_BB<-ggplot(gwas_data,aes(x=SNP,y=BB_Pvalue))+theme_bw()+geom_point(size =2)+
  geom_hline(yintercept=8, linetype="dashed", color = "red")+
  labs(x="",y="-log10 P-value") +custom_theme()

#############################################################
#
#
#
#					LD plot
#
#
#
###############################################################
WB_cov<-LD_list$WB
BB_cov<-LD_list$BB
melt_WB_cov<-matrix(ncol=3,nrow=0)
melt_WB_cov<-data.frame(melt_WB_cov)
melt_BB_cov<-matrix(ncol=3,nrow=0)
melt_BB_cov<-data.frame(melt_BB_cov)
for(SNP1 in 1:6){
  for(SNP2 in SNP1:6){
    melt_WB_cov<-rbind(melt_WB_cov,c(paste0("SNP",SNP1),paste0("SNP",SNP2),WB_cov[SNP1,SNP2]))
    melt_BB_cov<-rbind(melt_BB_cov,c(paste0("SNP",SNP1),paste0("SNP",SNP2),BB_cov[SNP1,SNP2]))
  }
}
colnames(melt_WB_cov)<-c("SNP1","SNP2","Value")
melt_WB_cov$Value<-(as.numeric(melt_WB_cov$Value))
colnames(melt_BB_cov)<-c("SNP1","SNP2","Value")
melt_BB_cov$Value<-(as.numeric(melt_BB_cov$Value))

WB_cor<-ggplot(melt_WB_cov, aes(x = SNP1, y = SNP2, fill = Value)) +
  geom_tile(aes( fill = abs(Value)))+scale_fill_stepsn(
    colors = c( "#F7F172","#F7CD72","#F7A972","#F78E72", "#F77772"),
    breaks = seq(0, 0.8, by = 0.2),
    limits = c(0, 1),
    show.limits = TRUE,
    na.value = 'grey50',
    name = "",guide = FALSE
  )+
  geom_text(aes(label=round(Value,2)),size = 5/14*5,fontface="bold")+
  theme_bw() +xlab("White British")+ylab("Correlation") +custom_theme()

BB_cor<-ggplot(melt_BB_cov, aes(x = SNP1, y = SNP2, fill = Value)) +
  geom_tile(aes( fill = abs(Value)))+scale_fill_stepsn(
    colors =  c( "#F7F172","#F7CD72","#F7A972","#F78E72", "#F77772"),
    breaks = seq(0, 0.8, by = 0.2),
    limits = c(0, 1),
    show.limits = TRUE,
    na.value = 'grey50',
    name = "",guide = FALSE
  )+
  geom_text(aes(label=round(Value,2)), size = 5/14*5, fontface="bold")+
  theme_bw() +xlab("Black British")+ylab("Correlation") +custom_theme()


#############################################################
#
#
#
#				Fine mapping plot
#
#
#
###############################################################
#############################################################
#					MESuSie 
#############################################################

MESuSiE_pip<-data.frame(MESuSiE_res$pip_config)
colnames(MESuSiE_pip)<-c("WB","BB","Shared")

Signal_vec <- MESuSiE_pip %>%
  mutate(Signal_vec = case_when(
    Shared > 0.5 ~ "Shared",
    WB > 0.5 ~ "White British",
    BB > 0.5 ~ "Black British",
    TRUE ~ "None"
  ))%>%pull(Signal_vec)

MESuSiE_plot_data<-data.frame(SNP=paste0("SNP",seq(1,6,1)),PIP=MESuSiE_res$pip,Signal =Signal_vec )
MESuSiE_plot_data$Signal<-factor(MESuSiE_plot_data$Signal,levels = c("White British","Black British","Shared","Either","None"))

MESuSiE_plot<-ggplot(MESuSiE_plot_data,aes(x=SNP,y=PIP,fill=Signal,shape = Signal))+theme_bw()+geom_point(size =3)+
  scale_fill_manual("Signal",values=c("White British"="#F2C1B6","Black British"="#ABDB9F","Shared" = "#6162B0","Either" = "#DB6C79","None"="gray"),drop = FALSE)+
  scale_shape_manual("Signal",values=c("White British"=24,"Black British" = 25,"Shared" = 23,"Either" = 22,"None"=20),drop = FALSE)+
  scale_y_continuous(limits=c(0, 1.1)) +geom_hline(yintercept=0.5, linetype="dashed")+
  labs(x="MESuSiE",y="PIP")                                                                                                                                                                                                                                                                           
MESuSiE_plot<-MESuSiE_plot+custom_theme()

#############################################################
#					SuSie WB
#############################################################
library(ggplot2)
library(cowplot)

SuSiE_pip<-data.frame(WB = susie_WB$pip, BB = susie_BB$pip)
colnames(SuSiE_pip)<-c("WB","BB")

Signal_vec <- SuSiE_pip %>%
  mutate(Signal_vec = case_when(
    WB > 0.5 & BB < 0.5 ~ "White British",
    BB > 0.5 & WB > 0.5~ "Shared",
    TRUE ~ "None"
  ))%>%pull(Signal_vec)

SuSiE_WB_data<-data.frame(SNP=paste0("SNP",seq(1,6,1)),PIP= susie_WB$pip,Signal =Signal_vec )
SuSiE_WB_data$Signal<-factor(SuSiE_WB_data$Signal,levels =  c("White British","Black British","Shared","Either","None"))

SuSiE_WB_plot<-ggplot(SuSiE_WB_data,aes(x=SNP,y=PIP,fill=Signal,shape = Signal))+theme_bw()+geom_point(size =3)+
  scale_fill_manual("Signal",values=c("White British"="#F2C1B6","Black British"="#ABDB9F","Shared" = "#6162B0","Either" = "#DB6C79","None"="gray"),drop = FALSE)+
  scale_shape_manual("Signal",values=c("White British"=24,"Black British" = 25,"Shared" = 23,"Either" = 22,"None"=20),drop = FALSE)+
  scale_y_continuous(limits=c(0, 1.1)) +geom_hline(yintercept=0.5, linetype="dashed")+
  labs(x="SuSiE White British",y="PIP")                                                                                                                                                                                                                                                                           
SuSiE_WB_plot<-SuSiE_WB_plot+custom_theme()
#############################################################
#					SuSie BB
###############################################################
Signal_vec <- SuSiE_pip %>%
  mutate(Signal_vec = case_when(
    WB < 0.5 & BB > 0.5 ~ "Black British",
    BB > 0.5 & WB > 0.5~ "Shared",
    TRUE ~ "None"
  ))%>%pull(Signal_vec)
SuSiE_BB_data<-data.frame(SNP=paste0("SNP",seq(1,6,1)),PIP= susie_BB$pip,Signal =Signal_vec )
SuSiE_BB_data$Signal<-factor(SuSiE_BB_data$Signal,levels = c("White British","Black British","Shared","Either","None"))

SuSiE_BB_plot<-ggplot(SuSiE_BB_data,aes(x=SNP,y=PIP,fill=Signal,shape = Signal))+theme_bw()+geom_point(size =3)+
  scale_fill_manual("Signal",values=c("White British"="#F2C1B6","Black British"="#ABDB9F","Shared" = "#6162B0","Either" = "#DB6C79","None"="gray"),drop = FALSE)+
  scale_shape_manual("Signal",values=c("White British"=24,"Black British" = 25,"Shared" = 23,"Either" = 22,"None"=20),drop = FALSE)+
    scale_y_continuous(limits=c(0, 1.1)) +geom_hline(yintercept=0.5, linetype="dashed")+
  labs(x="SuSiE Black British",y="PIP")                                                                                                                                                                                                                                                                           
SuSiE_BB_plot<-SuSiE_BB_plot+custom_theme()
#############################################################
#				Paintor
#############################################################

Paintor_pip<-data.frame( paintor_res$Posterior_Prob)
colnames(Paintor_pip)<-c("Either")

Signal_vec <- Paintor_pip %>%
  mutate(Signal_vec = case_when(
    Either > 0.5 ~ "Either",
    TRUE ~ "None"
  ))%>%pull(Signal_vec)

Paintor_plot_data<-data.frame(SNP=paste0("SNP",seq(1,6,1)),PIP=Paintor_pip$Either,Signal =Signal_vec )
Paintor_plot_data$Signal<-factor(Paintor_plot_data$Signal,levels =  c("White British","Black British","Shared","Either","None"))

Paintor_plot<-ggplot(Paintor_plot_data,aes(x=SNP,y=PIP,fill=Signal,shape = Signal))+theme_bw()+geom_point(size =3)+
  scale_fill_manual("Signal",values=c("White British"="#F2C1B6","Black British"="#ABDB9F","Shared" = "#6162B0","Either" = "#DB6C79","None"="gray"),drop = FALSE)+
  scale_shape_manual("Signal",values=c("White British"=24,"Black British" = 25,"Shared" = 23,"Either" = 22,"None"=20),drop = FALSE)+
  scale_y_continuous(limits=c(0, 1.1)) +geom_hline(yintercept=0.5, linetype="dashed")+
  labs(x="Paintor",y="PIP")                                                                                                                                                                                                                                                                           
Paintor_plot<-Paintor_plot+custom_theme()

library(ggpubr)
beta_gwas_test<-ggarrange(bar_WB,bar_BB,ncol=2,nrow=1)
locuszoom_test<-ggarrange(locuszoom_WB,locuszoom_BB,ncol=2,nrow=1)
LD_test<-ggarrange(WB_cor,BB_cor,ncol=2,nrow=1)
labeled_plot<-ggarrange(beta_gwas_test, locuszoom_test, LD_test,labels = c("a", "b", "c") ,font.label=list(color="black",size=7), nrow=3, ncol=1)
pip_plot<-ggarrange(SuSiE_WB_plot,SuSiE_BB_plot,MESuSiE_plot,Paintor_plot,ncol=2,nrow = 2 ,common.legend = TRUE, legend="bottom")
pip_plot<-ggarrange(pip_plot, labels = c("d"),font.label=list(color="black",size=7))
Fig_1<-ggarrange(labeled_plot,pip_plot,ncol=2,widths=c(4.5,5.5))
Fig_1
```

